---
title: 'Web Scraping Project: WTA Elo Ratings | Elena J. Muñoz'
output:
  html_document: default
---

## Project Summary

This project aims to demonstrate how to perform **Web Scraping in R** to extract **WTA (Women's Tennis Association) Elo rating data** from the site [Wheeloratings.com](https://www.wheeloratings.com/tennis_wta_ratings.html).\
The page displays data in an interactive table powered by `reactable.js`, so the values are not in a traditional HTML `<table>`, but embedded as a JSON object within a `<script>` element.

We will use the `rvest`, `jsonlite`, and `dplyr` packages to extract, convert, and visualize the data.

------------------------------------------------------------------------

## Step-by-step

```{r setup, include=FALSE}
# Load required packages
library(rvest)     # For reading and parsing HTML content
library(jsonlite)  # For interpreting and converting JSON into R objects
library(dplyr)     # For data manipulation
library(ggplot2)
```

```{r}
# Define the URL of the target web page
url <- "https://www.wheeloratings.com/tennis_wta_ratings.html"

# Read the HTML content of the page
page <- read_html(url)
```

```{r}
# Find the 'data-for' attribute of the <script> element that contains the JSON
# This identifies the widget containing the Elo ratings table
json_id <- page %>%
  html_elements("script[type='application/json']") %>%
  html_attr("data-for") %>%
  na.omit() %>%
  .[1]  # Select the first one, which contains the Elo table
```

```{r}
# Extract the JSON text from the relevant <script> block
json_text <- page %>%
  html_elements(paste0("script[data-for='", json_id, "']")) %>%
  html_text()

# Parse the JSON string into an R object
json_data <- fromJSON(json_text)

# Extract the data and convert it to a data.frame
elo_data <- as.data.frame(json_data$x$tag$attribs$data)
```


## Full WTA Elo Ratings Table

The following table presents the Elo ratings extracted directly from the website. It includes the Elo rank, player name, recent rank change, country, age, and surface-specific Elo scores.

```{r}
knitr::kable(elo_data, format = "html", table.attr = 'class="table table-striped"')
```

## Top 10 WTA Players by Elo Rating
Let's create a bar plot of the top 10 WTA players based on their overall Elo rating.

```{r}
# Select top 10 players by Elo rating
top10 <- elo_data %>%
  arrange(desc(elo)) %>%
  slice(1:10)

# Bar plot
ggplot(top10, aes(x = reorder(player_name, elo), y = elo)) +
  geom_col(fill = "#6a1b9a") +
  coord_flip() +
  labs(title = "Top 10 WTA Players by Elo Rating",
       x = "Player",
       y = "Elo Rating") +
  theme_minimal() +
  theme(text = element_text(size = 12))
```




  📫 Contact

📧 Email: elenamunoz412@gmail.com
